---
title: "cppRouting package"
author: "Vincent LARMET"
date: "15 juin 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'D:/users/vlarmet/Dijkstra/exemple/data')
```

#Package presentation

`cppRouting` is an `R` package which provide functions to calculate distances, shortest paths and isochrones/isodistances on non-negative weighted graphs. 
For now, `cppRouting` can implement :  
  
  - uni-directional Dijkstra algorithm,
  - bi-directional Dijkstra algorithm,
  - uni-directional A* algorithm  
  - New bi-directional A* algorithm (Piljs & Post, 2009 : see http://repub.eur.nl/pub/16100/ei2009-10.pdf)
   
All these functions are written in C++ and use std::priority_queue container from the Standard Template Library.  
This package have been made with `Rcpp` and `parallel` packages.

#Install
###Stable version from CRAN
```{r,eval=FALSE}
install.packages("cppRouting")
```
###or from github
```{r,eval=FALSE}
library(devtools)
devtools::install_github("vlarmet/cppRouting")
```
#Data
The data presented here is the official french road network describing over 500000 km of roads.  
All data used in this README are free and can be downloaded here :  
  
- roads :  http://professionnels.ign.fr/route500  
- general practitioners location : https://www.insee.fr/fr/statistiques/3568614?sommaire=3568656#consulter  
- maternity wards location : https://www.insee.fr/fr/statistiques/3568611?sommaire=3568656#dictionnaire  
- shapefile of the ~36000 communes in France  : http://professionnels.ign.fr/adminexpress  

Graph data have been preprocessed for more readability (see data_preparation.R).  

The final graph is composed of 234615 nodes and 685118 edges.  
Data has to be a 3 columns data.frame or matrix containing from, to and a cost/distance column. Here the cost is the time needed to travel in each edges (in minutes). From and to are vertices IDs (character or numeric).

#Main functions
  
`cppRouting` package provide these functions :  
  
- `get_distance_matrix` : compute distance matrix (between all combinations origin-destination nodes - *one-to-many*),  
- `get_distance_pair` : compute distances between origin and destination by pair (*one-to-one*),  
- `get_path_pair` : compute shortest paths between origin and destination by pair (*one-to-one*),  
- `get_multi_paths` : compute shortest paths between all origin nodes and all destination nodes (*one-to-many*),  
- `get_isochrone` : compute isochrones/isodistances with one or multiple breaks.    
                    
The choice between all the algorithms is available for *one-to-one* calculation like  `get_distance_pair` and `get_path_pair`.  
In these functions, uni-directional Dijkstra algorithm is stopped when the destination node is reached.  
`A*` and `NBA*` are relevant if geographic coordinates of all nodes are provided. Note that coordinates should be expressed in a **projection system**.  
To be accurate and efficient, `A*` and `NBA*` algorithms should use an admissible heuristic function (here the Euclidean distance), e.g cost and heuristic function must be expressed in the same unit.  
In `cppRouting`, heuristic function `h` is defined such that : h(xi,yi,xdestination,ydestination)/k, with a constant k; so in the case where coordinates are expressed in meters and cost is expressed in time, k is the maximum speed allowed on the road. By default, constant is 1 and is designed for graphs with cost expressed in the same unit than coordinates (for example in meters).  
  
If coordinates cannot be provided, bi-directional Dijkstra algorithm can offer a good alternative to A* in terms of performance.
  
##Examples
###Prepare data
```{r pressure, echo=TRUE,message=FALSE}
library(cppRouting)
library(dplyr)
library(sf)
library(ggplot2)
library(concaveman)
library(ggmap)

#Reading french road data
roads<-read.csv("roads.csv",colClasses = c("character","character","numeric"))
#Shapefile data of communes
com<-read_sf("com_simplified_geom.shp")
#Correspondance file between communes and nodes in the graph
ndcom<-read.csv("node_commune.csv",colClasses = c("character","character","numeric"))
#General practitioners locations
med<-read.csv("doctor.csv",colClasses = c("character","numeric","character","numeric"))
#Import nodes coordinates (projected in EPSG : 2154)
coord<-read.csv("coordinates.csv",colClasses = c("character","numeric","numeric"))
```

####Head of road network data
```{r , echo=TRUE,message=FALSE}
head(roads)
```

####Head of coordinates data  
```{r , echo=TRUE,message=FALSE}
head(coord)
```


###Instantiate the graph
```{r,echo=TRUE}
#Instantiate a graph with coordinates
graph<-makegraph(roads,directed = T,coords = coord)
```
###Distances by pairs between nodes (*one-to-one*)
####Using uni-directional Dijkstra algorithm
```{r,echo=TRUE}
#Generate 2000 random origin and destination nodes
origin<-sample(roads$from,2000)
destination<-sample(roads$from,2000)

#Uni-directional : single core
system.time(
pair_dijkstra<-get_distance_pair(graph,origin,destination)
)

```
####Using bi-directional Dijkstra algorithm
```{r,}
#Bi-directional : single core
system.time(
pair_bidijkstra<-get_distance_pair(graph,origin,destination,algorithm = "bi")
)

```
####Using A* algorithm
Coordinates are defined in meters and max speed is 110km/h; so for the heuristic function to be admissible, the constant equal 110/0.06 :  
```{r,echo=TRUE}
#A* single node
system.time(
pair_astar<-get_distance_pair(graph,origin,destination,algorithm = "A*",constant = 110/0.06)
)

```
####Using NBA* algorithm
```{r,echo=TRUE}
#NBA* single node
system.time(
pair_nba<-get_distance_pair(graph,origin,destination,algorithm = "NBA",constant = 110/0.06)
)

```
####Output
```{r,echo=TRUE}
head(cbind(pair_dijkstra,pair_bidijkstra,pair_astar,pair_nba))

```
#####In `get_distance_pair` function, all the algorithms can be ran in parallel by setting TRUE to allcores argument.  


###Compute isochrones
Let's compute isochrones around Dijon city  
```{r,echo=TRUE,message=FALSE,warning=FALSE}
#Compute isochrones
iso<-get_isochrone(graph,from = "205793",lim = c(15,25,45,60,90,120))
#Convert nodes to concave polygons with concaveman package
poly<-lapply(iso[[1]],function(x){
  x<-data.frame(noeuds=x,stringsAsFactors = F)
  x<-left_join(x,coord,by=c("noeuds"="ID"))
  return(concaveman(summarise(st_as_sf(x,coords=c("X","Y"),crs=2154))))
})

poly<-do.call(rbind,poly)
poly$time<-as.factor(names(iso[[1]]))
#Multipolygon
poly2<-st_cast(poly,"MULTIPOLYGON")
poly2$time<-reorder(poly2$time,c(120,90,60,45,25,15))
#Reproject for plotting
poly2<-st_transform(poly2,"+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")
#Import map backgroung
dijon=get_map(location=c(lon=5.041140,lat=47.323025),zoom=7, source="google",maptype = "toner-2010")
#Plot the map
p<-ggmap(dijon)+
  geom_sf(data=poly2,aes(fill=time),alpha=.8,inherit.aes = FALSE)+
  scale_fill_brewer(palette = "YlOrRd")+
  labs(fill="Minutes")+
  ggtitle("Isochrones around Dijon")+
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.y=element_blank(),axis.title.x=element_blank())
p

```


#Applications
## Application 1 : Calculate Two Step Floating Catchment Areas (2SFCA) of general practitioners in France

2SFCA method is explained here : https://en.wikipedia.org/wiki/Two-step_floating_catchment_area_method  


###First step
Isochrones are calculated with the `cppRouting` function `get_isochrone` 
```{r,echo=TRUE}
#Isochrone around doctor locations with time limit of 15 minutes
iso<-get_isochrone(graph,from = ndcom[ndcom$com %in% med$CODGEO,"id_noeud"],lim = 15)
#Convert list to long data frame
df<-stack(setNames(iso, seq_along(iso)))
df$ind<-rep(names(iso),times=sapply(iso,length))
df<-df[df$values %in% ndcom$id_noeud,]
#Joining and summing population located in each isochrone
df<-left_join(df,ndcom[,c("id_noeud","POPULATION")],by=c("values"="id_noeud"))
df<-df %>% group_by(ind) %>%
  summarise(pop=sum(POPULATION))
#Joining number of doctors 
df<-left_join(df,med[,c("id_noeud","NB_D201")],by=c("ind"="id_noeud"))
#Calculate ratios
df$ratio<-df$NB_D201/df$pop
```

###Second step
```{r,echo=TRUE,message=FALSE}
#Isochrone around each commune with time limit of 15 minutes (few seconds to compute)
iso2<-get_isochrone(graph,from=ndcom$id_noeud,lim = 15)
#Convert list to long data frame
df2<-stack(setNames(iso2, seq_along(iso2)))
df2$ind<-rep(names(iso2),times=sapply(iso2,length))
#Joining and summing ratios calculated in first step
df2<-left_join(df2,df[,c("ind","ratio")],by=c("values"="ind"))
df2<-df2 %>% group_by(ind) %>%
  summarise(sfca=sum(ratio,na.rm=T))
```

###Plot the map for Bourgogne-Franche-Comte region
```{r,echo=TRUE,message=FALSE}
#Joining commune IDs to nodes
df2<-left_join(df2,ndcom[,c("id_noeud","com")],by=c("ind"="id_noeud"))
#Joining 2SFCA to shapefile
com<-left_join(com,df2[,c("com","sfca")],by=c("INSEE_COM"="com"))
#Plot for one region
p<-ggplot()+
  geom_sf(data=com[com$NOM_REG=="BOURGOGNE-FRANCHE-COMTE",],aes(fill=sfca),colour=NA)+
  coord_sf(datum=NA)+
  scale_fill_gradient(low="#BB2528",high = "#FFFF66")+
  labs(fill="2SFCA")+
  ggtitle("2SFCA applied to general practitioners")
p
```

##Application 2 : Calculate the minimum travel time to the closest maternity ward in France  
```{r,echo=TRUE,message=FALSE}
#Import materinty ward locations
maternity<-read.csv("maternity.csv",colClasses = c("character","numeric"))
```

###Shortest travel time matrix
The shortest travel time is computed with the `cppRouting` function `get_distance_matrix`. 
In order to compute multiple distances from one source, original uni-directional Dijkstra algorithm is ran without early stopping.  
We compute travel time from all commune nodes to all maternity ward nodes (e.g ~36000*400 distances).
```{r,echo=TRUE,message=FALSE,warning=FALSE,include=TRUE}
#Distance matrix (around 10 minutes to compute)
dists<-get_distance_matrix(graph,
                           from=ndcom$id_noeud,
                           to=ndcom$id_noeud[ndcom$com %in% maternity$CODGEO],
                           allcores=TRUE)
#We extract each minimum travel time for all the communes
dists2<-data.frame(node=ndcom$id_noeud,mindist=apply(dists,1,min,na.rm=T))
#Joining commune IDs to nodes
dists2<-left_join(dists2,ndcom[,c("id_noeud","com")],by=c("node"="id_noeud"))
#Joining minimum travel time to the shapefile
com<-left_join(com,dists2[,c("com","mindist")],by=c("INSEE_COM"="com"))
```

#Plot the map of minimum travel time in Bourgogne-Franche-Comte region
```{r,echo=TRUE,message=FALSE}
p<-ggplot()+
  geom_sf(data=com[com$NOM_REG=="BOURGOGNE-FRANCHE-COMTE",],aes(fill=mindist),colour=NA)+
  coord_sf(datum=NA)+
  scale_fill_gradient(low="#009900",high="#003300")+
  labs(fill="Minutes")+
  ggtitle("Travel time to the closest maternity ward")
p
```


#Benchmark with other R packages
To show the efficiency of `cppRouting`, we can make some benchmarking with the famous R package `igraph`, and the `dodgr` package which provide highly optimized heaps. 

###Distance matrix : one core
```{r,echo=TRUE,warning=FALSE,message=FALSE}
library(igraph)
library(dodgr)
#Sampling 1000 random origin/destination nodes (1000000 distances to compute)
origin<-sample(unique(roads$from),1000,replace = F)
destination<-sample(unique(roads$from),1000,replace = F)
```

```{r, echo=TRUE,warning=FALSE}
#igraph 
graph_igraph<-graph_from_data_frame(roads,directed = TRUE)

system.time(
  test_igraph<-distances(graph_igraph,origin,to=destination,weights = E(graph_igraph)$weight,mode="out")
)


#dodgr
#Adding coordinates to data
roads2<-roads
colnames(roads2)[3]<-"dist"
roads2<-left_join(roads2,coord,by=c("from"="ID"))
colnames(roads2)[4:5]<-c("from_lon","from_lat")
roads2<-left_join(roads2,coord,by=c("to"="ID"))
colnames(roads2)[6:7]<-c("to_lon","to_lat")
colnames(roads2)[1:2]<-c("from_id","to_id")
roads2$from_id<-as.character(roads2$from_id)
roads2$to_id<-as.character(roads2$to_id)

system.time(
test_dodgr<-dodgr_dists(graph=data.frame(roads2),from=origin,to=destination,parallel=FALSE)
)

#cppRouting
system.time(
test_cpp<-get_distance_matrix(graph,origin,destination,allcores = FALSE)
)
```
####Ouput
```{r,echo=TRUE}
head(cbind(test_igraph[,1],test_dodgr[,1],test_cpp[,1]))
```

###Distance matrix : parallel
```{r,echo=TRUE,warning=FALSE}
#dodgr
system.time(
test_dodgr<-dodgr_dists(graph=data.frame(roads2),from=origin,to=destination,parallel=TRUE)
)

#cppRouting
system.time(
test_cpp<-get_distance_matrix(graph,origin,destination,allcores = TRUE)
)
```

##Benchmark on computing shortest paths by pairs
```{r,echo=TRUE,warning=FALSE}
#Sampling 500 random origin/destination nodes 
origin<-sample(unique(roads$from),500,replace = F)
destination<-sample(unique(roads$from),500,replace = F)
#dodgr
system.time(
test_dodgr<-dodgr_paths(graph=data.frame(roads2),from=origin,to=destination,pairwise = TRUE)
)

#cppRouting
system.time(
test_cpp<-get_path_pair(graph,origin,destination,algorithm = "NBA",constant=110/0.06)
)
```
###Test similarity of the first travel
```{r,echo=TRUE,warning=FALSE}
#Number of nodes
length(test_dodgr[[1]][[1]])
length(test_cpp[[1]])

#Setdiff 
setdiff(test_dodgr[[1]][[1]],test_cpp[[1]])

```

#New algorithms `cppRouting` will provide in the future   
  
- Detours admitting shortest paths : finding the nodes that are reachable under a fixed detour time around the shortest path  
- Graph simplification by removing irrelevant nodes in order to compute in a faster way the  shortest distance or travel time  
- Contraction hierarchies implementation  